Luminance constancy under fixed geometry

### 2018-03-26 Draft2/draft2.tex
Today’s edits

\title{Luminance constancy under fixed geometry}
\abstract{The light that reflects off an object surface and is captured by our eyes varies significantly with its context. But, the human visual system has the ability to stably perceive the object's color invariant of the context variations. The mechanism of this behaviorally significant invariance detection ability remains largely unknown. To understand the computational principles that lead to such invariant perception, we study the perception of object luminance under spectral variations in 3D scenes. Specifically, we study the effects of variations in reflectance and illumination spectra in a scene on the perceived luminance of an object. We have developed a software that generates naturalistic multispectral images of 3D scenes with precise control over the geometrical and spectral properties of the constituents that make a scene. We label these images with the luminance of a specific object in the scene. Next, we simulate the response of the retinal cones to these multispectral images using an accurate model of the early visual system. We then use supervised learning methods on these labeled cone responses to identify the computations that lead to accurate luminance estimation. We show that if only either target, or background, or illumination spectra are allowed to vary, the standard luminance can be estimated through simple transformations of the cone responses. When all the spectra vary simultaneously, while it is not easy to recover the luminance through simple transformations, a decoding scheme that compares the light from the target and the surround, and properly adds the response of the L,M and S cones can recover the luminance within about 15\% RMSE.}

@@

% Figure 1: Introduction
\begin{figure}
\centering
\begin{subfigure}{0.4 \textwidth}
		\centering
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_a.png}
        \caption{}
        \label{fig:introSchematic}
    \end{subfigure}
    \begin{subfigure}{0.55 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_b.png}
        \caption{}
        \label{fig:introExampleFigure}
    \end{subfigure}
    \label{introFigure}
    \caption{(a) {\bf Color Constancy:} The light reflected from an object to the eye depends both on the object's surface reflectance of the object and the illumination. It also depends on geometric factors, such as the object's shape, pose, and position relative to the observer. The human visual system is able to account for variations in the reflected light due to object-extrinsic factors and produce a percept of object color that is relatively stable, an ability called color constancy. (b) {\bf Luminance Constancy:} Images of a sphere under a fixed illuminant.  Down each column, the reflectance function of the sphere varies only by a scale factor; the relative reflectance spectrum held fixed.  The relative spectrum (shape of reflectance spectrum) varies across each row.
Within each row, the object luminance of the spheres is constant; we define object luminance as the luminance of the light that would be reflected by the object's surface under standard illuminant CIE D65.  The values on the left of each row provide the corresponding object luminance. We cast the problem of computational luminance constancy to be that of estimating object luminance from the image, across variation in other scene factors: variation in relative surface reflectance, variation in illumination, and variation in the reflectance of the other objects in the scene.}
 \end{figure}

@@

We define the computational problem of luminance constancy as that of estimating the luminous reflectance factor of the diffuse component of an object's surface reflectance function.
The luminous reflectance factor is a measure of the overall amount of light reflected by a matte surface.
More precisely, the luminous reflectance factor is the normalized luminance of the light that would
be reflected to the eye from a flat matte object with that the specified surface reflectance function, when the object is diffusely illuminated by a reference illuminant.
The normalization factor is 100 divided by the luminance of the reference illuminant, so that luminous reflectance factors range between 0 and 100.
Here we use CIE daylight D65 as the reference illuminant and the CIE 1931 photopic luminosity function to compute luminance  \cite{CIE86}.

% MOVE THIS TO DICUSSION
%This general approach has been adopted by other labs for the study of computational estimation of optical flow (Baker et al., 2011) and is becoming increasingly popular (Kovacs, Bell, Snavely, & Bala, 2017), as well as for the generation of stimuli for the study of human perception (e.g., Boyaci, Maloney, & Hersh, 2003; Todd & Mingolla, 1983; Arend & Reeves, 1986; Johnston, Cumming, & Parker, 1993). Of course, success within the domain of graphics images does not guarantee smooth generalization to real natural images, a point we return to in the Discussion.
% \cite{baker2011database}  \cite{kovacs17shading} \citeNP{boyaci2003effect, todd1983perception, arend1986Simultaneous, johnston1993integration}
% Autonmous vehicle work - ask BAW

% NOT SURE WE NEED ANY OF THIS.  PROBABLY NOT HERE IF WE DO
%Even within the special case of luminance constancy, there are many scene manipulations that could be considered. The set of manipulations we focus on here are illustrated by Fig.~\ref{fig:introExampleFigure} and the three panels of Fig.~\ref{fig:studiedCases}. These cases are: 1. variation in target object spectrum with fixed light source and background objects spectra (Fig.~\ref{fig:introExampleFigure}), 2. variation in target object spectrum and background objects spectra, with fixed light source  (Fig.~\ref{fig:backgroundVarying}), 3. variations in both the target object spectrum and the light source spectra, with fixed background object spectra (Fig.~\ref{fig:targetIlluminantVarying}), and 3. variations in all three types of spectra (target object, light source, and background objects) (Fig.~\ref{fig:allSpectraVarying}). Across real scenes, there will also be variations in the shape of the target object, its pose, eye position, etc. (see \nameref{Methods}). In this paper, we focus on the problem of luminance constancy within scene ensembles where these geometric factors are held fixed.
%
%We show that for cases that have spectral variation only either in target object surface reflectance or background object surface reflectance or the power spectrum of the light sources, luminance can be estimated by simple transformations of the cone responses. For variations only in target object reflectance spectra (case 1) or just the target and background objects reflectance spectra (case 2), object luminance can be recovered directly from the light reflected off the target object. For case 3, where both the target object reflectance and light source spectra vary, object luminance can be recovered from the contrast between the target and the background. When all the spectral features in the scene are allowed to change (case 4), the standard luminance can not be recovered simply from the cone response or the contrast. We have used AMA with a Bayesian estimator to estimate standard luminance of the target object. Using this method, we can recover the luminance of the target object with $\sim 15\%$ relative root mean square error.  Our analysis shows that the optimal receptive fields for estimating the luminance of the target object show a center surround structure, supporting a comparison between the light coming from the target object and its surround. Additionally, luminance estimation involves a weighted sum of the L, M and S cone responses where the receptive fields give more weight to the L and M cone responses compared to S cone responses. Comparing the performance of the RFs learnt using  one case on the cone response of other cases shows that the receptive field of the complex cases perform well on images of simpler cases. We discuss the implications of such receptive fields on luminance constancy.

@@

In the present work, we used our package to generate datasets of naturalistic scenes and corresponding images.
We used one fixed base scene and inserted a target object of fixed shape and size into this scene.
Across scenes, we systematically varied the reflectance of the target object, and randomly varied the reflectance of the other objects and the spectral power distributions of the light sources.

@@

\subsubsection*{Accuracy maximization analysis}
We perform two steps to decode the luminance of the target object from the cone isomerization. First, we use AMA to learn the linear filters that are optimal for estimating the target object luminance. Second, we determine how to optimally decode the luminance from the response of these filters to arbitrary stimuli.

@@

More concretely, let us assume that there are $N_{\nu}$ possible values of the target object luminance levels ($\{Y_k: k\in[1,N_{\nu}] \}$) and the $l$'th image corresponding to the $k$'th luminance level is represented as $s_{kl}$. Also, assume that there is a set of linear receptive fields {\textbf{\textit f}} which when acting on $s_{kl}$ gives the response ${\textbf{\textit R}}_{kl}$. Also assume there is a decoder $g$, with the knowledge of the response model and of every stimulus in the training set, that returns the Bayes' optimal estimate of the luminance level ($\hat{Y}_k$) of the target object in each stimulus $s_{kl}$. The crux computation for obtaining this estimate is the computation of the posterior probability of the luminance given an observed response, $P(Y|\textbf{R}_{kl})$ \cite{geisler2009optimal,burge2017accuracy,jaini2017linking}
. With this luminance estimate, we calculate the cost given the cost function $\mathcal{C}_{kl} = \gamma(Y_{kl},\hat{Y}_{kl})$. Finally, AMA searches for the set of optimal receptive fields ${\textbf{\textit f}}^{\rm opt}$ that minimize the cost over the entire training set $\{s_{kl}: k\in[1,N_{\nu}], l\in[1,N_k]\}$ where $N_{k}$ is the number of stimuli corresponding to the label $Y_k$. Thus,
\begin{align}
{\textbf {\textit f} }^{\rm opt} = \argmin_{\textbf {\textit f} } \sum_{k=1}^{N_{\nu}}  \sum_{l=1}^{N_k} {\mathcal{C}}_{kl}.
\label{eq:fopt}
\end{align}

@@

\section{Results} \label{Results}
In this section, we discuss the results of supervised learning to estimate the luminance of the target object on the the various cases we described previously. We describe the computations that need to be performed on the cone photoreceptor responses to estimate the luminance. We start with case 1, where only the reflectance spectrum of the target object is allowed to vary, and show that in this case the luminance can be estimated directly from the photoreceptor response. We then progressively increase the complexity of the spectral conditions and explain the correspondingly complex computations that are required to estimate the target object luminance. Next, we show the optimal AMA receptive fields for each one of the spectral cases discussed in this work. We show that the optimal receptive fields for target luminance estimation have a center surround structure and they give more weights to the response of L and M cones than S cones. We finally show that, as expected, the receptive fields for the complex spectral cases generalize to the simple cases. Hence, one requires only one set of optimal receptive fields for estimating the target luminance, not separate receptive fields for each spectral case.

\subsection{Luminance estimates for each spectral case}
\subsubsection{Case 1: Target object relative surface reflectance spectrum variable, light source and background object spectra fixed}
We start with the results of the simplest type of spectral variation where only the target object relative surface reflectance spectrum is allowed to change. 
We used AMA to learn a set of linear receptive fields that are optimal for estimating target object luminance from the model cone photoreceptor response to the images from this case.
Fig.~\ref{fig:case1RFResponse} shows the projection of the photoreceptor response along the first two optimal AMA receptive fields. 
Clearly, there exists a subspace along which the photoreceptor response separates according to the luminance level assigned to the target object. 
Thus, the target object luminance can be obtained by projecting the photoreceptor response along these AMA receptive fields. 
We used such projections and a generalized Bayes' decoder (see methods) to estimate the luminance of the target objects in the test set stimuli. 
Fig.~\ref{fig:case9Results} shows the estimated luminance v/s the actual luminance assigned to the target object for the test stimuli in case 1. 
We show the estimates obtained by AMA and two null models: a naive method that assigns the average target object luminance across the training set to all images, and a linear model based on the photoreceptors corresponding to the center pixel on the target object (see Methods). 
As expected, in the absence of any external source of variation, the photoreceptor response is proportional to the luminance of the target object, and hence can be easily estimated by both AMA and linear regression.

@@

\subsubsection{Case 2: Target object relative surface reflectance spectrum variable, light source spectra fixed, background object spectra variable}
When the light source emission spectrum is fixed, the naive expectation would be that the luminance of the target object would depend only on the surface reflectance of the target object. 
Indeed, it is believed that the variation in the surface reflectance of the background has little effect on the reflected light from the target object  \cite{BrainardWandellRetinex}. 
But our analysis shows that the secondary reflections from the target object, of the light coming indirectly after being reflected from the background objects, have a significant effect on the response of the photoreceptors corresponding to the target object. 
Fig.~{\ref{fig:case2RFResponse}} shows the projection of the photoreceptor response for the stimuli in case 2 along the first two AMA receptive fields. 
These receptive fields were learnt using the stimuli from the same case 2. 
The response projected along the first two optimal receptive fields do not separate as well compared to case 1. 
But the overlap is lower if the projection is considered in subspaces that have more AMA receptive fields. 
Fig.~\ref{fig:ErrorVsNFilters} shows the relative error of luminance estimate as a function of the number of AMA receptive fields used for luminance estimation. 
As expected, the relative error decreases with the number of receptive fields and converges to the minimum value around 8 receptive fields. 
Fig.~\ref{fig:case2Estimates} shows the luminance estimates for the learning methods used in this work. 
The mean estimates fall along the diagonal x=y line. 
The variance in the estimates increases with assigned luminance. 
This is because the isomerization noise is directly proportional to the light reflected from the target object and is higher at higher values of the target object luminance.

%% Figure Case 2
\begin{figure}
\centering
    \begin{subfigure}[b]{0.3	 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure10/Figure10_a.pdf}
        \caption{RF response Case 2}
        \label{fig:case2RFResponse}
    \end{subfigure}
        \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure10/Figure10_b.pdf}
        \caption{Luminance Estimates}
        \label{fig:case2Estimates}
    \end{subfigure}
    \caption{{\bf Case 2:} (a) Photoreceptor response projected along the first two AMA receptive fields. (b) Estimated v/s assigned target object luminance for the image in case 2. Solid lines show the mean estimate, the filled region in light color  shows one standard deviation.}
\label{fig:case2Results}
\end{figure}

@@

\subsubsection{Case 3: Target object relative surface reflectance spectrum variable, light source spectra variable, background object spectra fixed}
For the spectral variations of case 3, the AMA receptive fields learnt on the photoreceptor response to the images does a very poor job in predicting the target object luminance (Fig.~\ref{fig:isomerizationFails}). 
This is because, since the illuminant spectra varies in this case, the photoreceptor response is not directly proportional to the target object surface anymore. 
But, since the background surfaces are fixed in this case, the ratio between the response of  photoreceptors corresponding to the target object and the background remains proportional to the target object luminance. 
This is because a change in illuminant spectra produces the same effect in both the target object and the background object. 
Thus, the AMA receptive fields learnt on the contrast normalized photoreceptor response (see Methods) can be used to predict the target object luminance (Fig.~\ref{fig:constrastWorks}). 
Fig.~\ref{fig:case10Results} shows the estimated target object luminance v/s the assigned luminance for the images in case 3 obtained using the contrast normalized representation. 
AMA receptive fields provide excellent predictions of the target object luminance. 
The predictions are much better compared to that of a linear model that only considers the (contrast normalized) response of the photoreceptors corresponding to the center pixel of the target object.

%% Figure Case 3: Isomerization Fails
\begin{figure}
\centering
    \begin{subfigure}[b]{0.3	 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure12/Figure12_a.pdf}
        \caption{RFs learnt using photoreceptor response}
        \label{fig:isomerizationFails}
    \end{subfigure}
        \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure12/Figure12_b.pdf}
        \caption{RFs learnt using contrast signal}
        \label{fig:constrastWorks}
    \end{subfigure}
            \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure12/Figure12_c.pdf}
        \caption{Luminance estimates}
        \label{fig:case10Results}
    \end{subfigure}
    \caption{{\bf Case 3: Luminance can be estimated using the contrast signal:} (a) Photoreceptor response projected on AMA receptive fields calculated using photoreceptor response. (b) The contrast normalized photoreceptor response projected along the first two AMA receptive fields. The receptive fields were calculated using the contrast normalized stimuli. The response at each luminance level separates much better in this representation. (c) Estimated v/s assigned target object luminance for the image in case 3. Solid lines show the mean estimate, the filled region in light color  shows one standard deviation. AMA estimates are much better compared to the null models.}
\label{fig:importanceOfConstrast}
\end{figure}

@@

\subsubsection{Case 4: Target object relative surface reflectance spectrum variable, light source spectra variable, background object spectra variable}
% Figure Case 4
\begin{figure}
\centering
    \begin{subfigure}[b]{0.3 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure13/Figure13_a.pdf}
        \caption{AMA RF Response}
        \label{fig:case12FiltersResponse}
    \end{subfigure}    
        \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure13/Figure13_b.pdf}
        \caption{Luminance Estimates}
        \label{fig:case12Results}
    \end{subfigure}
    \caption{{\bf Case 3:} (a) Contrast normalized photoreceptor response projected along the first two AMA receptive fields. (b) Target object luminance estimates v/s assigned luminance for images in case 3. The AMA estimates are unbiased and have smaller relative error compared to linear model.}
    \label{fig:case12AllResults}
\end{figure}

Case 4 is the most complex type of variation that we consider in this work. Here we allow all three types of spectra to change from image to image. In this case, neither the photoreceptor response (as in Case 1 and 2), nor any simple transformation of the response (case 3) can provide the target object luminance.We used AMA to learn the receptive fields using the contrast normalized cone photoreceptor response from stimuli in case 4. Fig.~\ref{fig:case12AllResults} shows the AMA receptive field response and the luminance estimates for the images in case 4. 
We have used the first 8 receptive fields to estimate the luminance in this case (see Fig.\ref{fig:ErrorVsNFilters}). 
The AMA receptive fields do an excellent job of predicting the target object luminance. 
The estimates are unbiased with the mean estimates lying close the diagonal unit slope line (Fig.~\ref{fig:case12Results}). 
The relative error of estimation is $\sim15\%$, which low compared to that of the naive model ($\sim42\%$) or the linear model on center pixel ($\sim21\%$).

>>>>

\title{Luminance constancy under fixed geometry}
\abstract{The light that reflects off an object surface and is captured by our eyes varies significantly with its context. But, the human visual system has the ability to stably perceive the object's color invariant of the context variations. The mechanism of this behaviorally significant invariance detection ability remains largely unknown. To understand the computational principles that lead to such invariant perception, we study the perception of LRV under spectral variations in 3D scenes. Specifically, we study the effects of variations in reflectance and illumination spectra in a scene on the perceived luminance of an object. We have developed a software that generates naturalistic multispectral images of 3D scenes with precise control over the geometrical and spectral properties of the constituents that make a scene. We label these images with the luminance of a specific object in the scene. Next, we simulate the response of the retinal cones to these multispectral images using an accurate model of the early visual system. We then use supervised learning methods on these labeled cone responses to identify the computations that lead to accurate luminance estimation. We show that if only either target, or background, or illumination spectra are allowed to vary, the standard luminance can be estimated through simple transformations of the cone responses. When all the spectra vary simultaneously, while it is not easy to recover the luminance through simple transformations, a decoding scheme that compares the light from the target and the surround, and properly adds the response of the L,M and S cones can recover the luminance within about 15\% RMSE.}

@@

% Figure 1: Introduction
\begin{figure}
\centering
\begin{subfigure}{0.4 \textwidth}
		\centering
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_a.png}
        \caption{}
        \label{fig:introSchematic}
    \end{subfigure}
    \begin{subfigure}{0.55 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure1/Figure1_b.png}
        \caption{}
        \label{fig:introExampleFigure}
    \end{subfigure}
    \label{introFigure}
    \caption{(a) {\bf Color Constancy:} The light reflected from an object to the eye depends both on the object's surface reflectance of the object and the illumination. It also depends on geometric factors, such as the object's shape, pose, and position relative to the observer. The human visual system is able to account for variations in the reflected light due to object-extrinsic factors and produce a percept of object color that is relatively stable, an ability called color constancy. (b) {\bf Luminance Constancy:} Images of a sphere under a fixed illuminant.  Down each column, the reflectance function of the sphere varies only by a scale factor; the relative reflectance spectrum held fixed.  The relative spectrum (shape of reflectance spectrum) varies across each row.
Within each row, the light reflectance value (LRV) of the spheres is constant. The values on the left of each row provide the corresponding LRV. We cast the problem of computational luminance constancy to be that of estimating LRV from the image, across variation in other scene factors: variation in relative surface reflectance, variation in illumination, and variation in the reflectance of the other objects in the scene.}
 \end{figure}

@@

% I found the definition of light reflectance value on the web, sort of.  Need to find the right citable reference and or a term for the same thing that
% we can find a reference for.
We define the computational problem of luminance constancy as that of estimating the light reflectance value (LRV) of an object's surface reflectance function.
The LRV is a measure of the overall amount of light reflected by surface.
More precisely, the LRV is the luminance of the light that would
be reflected to the eye from an object with the specified surface reflectance function when the object is illuminated by a reference illuminant,
normalized by the luminance of the reference illuminant itself.
Here we use CIE daylight D65 as the reference illuminant and the CIE 1931 photopic luminosity function \cite{CIE86}.
LRVs are typically expressed in percent, so that they range from 0 to 100.

% MOVE THIS TO DISCUSSION
%This general approach has been adopted by other labs for the study of computational estimation of optical flow (Baker et al., 2011) and is becoming increasingly popular (Kovacs, Bell, Snavely, & Bala, 2017), as well as for the generation of stimuli for the study of human perception (e.g., Boyaci, Maloney, & Hersh, 2003; Todd & Mingolla, 1983; Arend & Reeves, 1986; Johnston, Cumming, & Parker, 1993). Of course, success within the domain of graphics images does not guarantee smooth generalization to real natural images, a point we return to in the Discussion.
% \cite{baker2011database}  \cite{kovacs17shading} \citeNP{boyaci2003effect, todd1983perception, arend1986Simultaneous, johnston1993integration}
% Autonmous vehicle work - ask BAW

% NOT SURE WE NEED ANY OF THIS.  PROBABLY NOT HERE IF WE DO
%Even within the special case of luminance constancy, there are many scene manipulations that could be considered. The set of manipulations we focus on here are illustrated by Fig.~\ref{fig:introExampleFigure} and the three panels of Fig.~\ref{fig:studiedCases}. These cases are: 1. variation in target object spectrum with fixed light source and background objects spectra (Fig.~\ref{fig:introExampleFigure}), 2. variation in target object spectrum and background objects spectra, with fixed light source  (Fig.~\ref{fig:backgroundVarying}), 3. variations in both the target object spectrum and the light source spectra, with fixed background object spectra (Fig.~\ref{fig:targetIlluminantVarying}), and 3. variations in all three types of spectra (target object, light source, and background objects) (Fig.~\ref{fig:allSpectraVarying}). Across real scenes, there will also be variations in the shape of the target object, its pose, eye position, etc. (see \nameref{Methods}). In this paper, we focus on the problem of luminance constancy within scene ensembles where these geometric factors are held fixed.
%
%We show that for cases that have spectral variation only either in target object surface reflectance or background object surface reflectance or the power spectrum of the light sources, luminance can be estimated by simple transformations of the cone responses. For variations only in target object reflectance spectra (case 1) or just the target and background objects reflectance spectra (case 2), LRV can be recovered directly from the light reflected off the target object. For case 3, where both the target object reflectance and light source spectra vary, LRV can be recovered from the contrast between the target and the background. When all the spectral features in the scene are allowed to change (case 4), the standard luminance can not be recovered simply from the cone response or the contrast. We have used AMA with a Bayesian estimator to estimate standard luminance of the target object. Using this method, we can recover the luminance of the target object with $\sim 15\%$ relative root mean square error.  Our analysis shows that the optimal receptive fields for estimating the luminance of the target object show a center surround structure, supporting a comparison between the light coming from the target object and its surround. Additionally, luminance estimation involves a weighted sum of the L, M and S cone responses where the receptive fields give more weight to the L and M cone responses compared to S cone responses. Comparing the performance of the RFs learnt using  one case on the cone response of other cases shows that the receptive field of the complex cases perform well on images of simpler cases. We discuss the implications of such receptive fields on luminance constancy.

@@

In the present work, we used our package to generate datasets of naturalistic scenes and corresponding images.
We used one fixed base scene and inserted a target object of fixed shape and size into this scene.
Across scenes, we systematically varied the reflectance of the target object, and randomly varied the reflectance of the other objects and the spectral power distributions of the light sources.
We generated five distinct datasets.
Across these datasets, the luminance constancy problem is progressively more difficult (Fig.~\ref{fig:studiedCases}).
In the first dataset (Condition 0), the only source of stimulus variation was in the task relevant quantity, the LRV.
Here, we fixed the relative reflectance spectrum of the target object, the reflectance spectra of the other objects in the scene, and the spectral power distribution of the light sources fixed.
We used 10 LRV values, equally spaced between 20 and 60.
Condition 0 provides a baseline.

In the remaining conditions, we systematically introduced variation in scene factors other than the LRV.

In the second dataset (Condition 1), we introduced variation in the relative reflectance spectrum of the target object.
In essence, this condition introduces color variation across target objects that have the same LRV.

Conditions 2 and 3 each increase the variation relative to Condition 1.

In Condition 2, in addition to the relative reflectance spectrum of the target
we varied the reflectance spectra of the background objects,
while holding the the spectral power distribution of the light sources
constant.

In Condition 3, in addition to the relative reflectance spectrum of the target
we varied the spectral power distribution of the light sources,
while holding the reflectance spectra of the background objects constant.

Finally, in Condition 4 we varied all three factors (relative reflectance spectrum,
reflectance spectra of the background objects, spectral power distribution of the light sources).

Condition 0: Baseline
Condition 1: Target Color Only
Condition 2: Target Color and Background

@@

\subsubsection*{Accuracy maximization analysis}
We perform two steps to decode the luminance of the target object from the cone isomerization. First, we use AMA to learn the linear filters that are optimal for estimating the target LRV. Second, we determine how to optimally decode the luminance from the response of these filters to arbitrary stimuli.

@@

More concretely, let us assume that there are $N_{\nu}$ possible values of the target LRV levels ($\{Y_k: k\in[1,N_{\nu}] \}$) and the $l$'th image corresponding to the $k$'th luminance level is represented as $s_{kl}$. Also, assume that there is a set of linear receptive fields {\textbf{\textit f}} which when acting on $s_{kl}$ gives the response ${\textbf{\textit R}}_{kl}$. Also assume there is a decoder $g$, with the knowledge of the response model and of every stimulus in the training set, that returns the Bayes' optimal estimate of the luminance level ($\hat{Y}_k$) of the target object in each stimulus $s_{kl}$. The crux computation for obtaining this estimate is the computation of the posterior probability of the luminance given an observed response, $P(Y|\textbf{R}_{kl})$ \cite{geisler2009optimal,burge2017accuracy,jaini2017linking}
. With this luminance estimate, we calculate the cost given the cost function $\mathcal{C}_{kl} = \gamma(Y_{kl},\hat{Y}_{kl})$. Finally, AMA searches for the set of optimal receptive fields ${\textbf{\textit f}}^{\rm opt}$ that minimize the cost over the entire training set $\{s_{kl}: k\in[1,N_{\nu}], l\in[1,N_k]\}$ where $N_{k}$ is the number of stimuli corresponding to the label $Y_k$. Thus,
\begin{align}
{\textbf {\textit f} }^{\rm opt} = \argmin_{\textbf {\textit f} } \sum_{k=1}^{N_{\nu}}  \sum_{l=1}^{N_k} {\mathcal{C}}_{kl}.
\label{eq:fopt}
\end{align}

@@

\section{Results} \label{Results}
In this section, we discuss the results of supervised learning to estimate the luminance of the target object on the the various cases we described previously. We describe the computations that need to be performed on the cone photoreceptor responses to estimate the luminance. We start with case 1, where only the reflectance spectrum of the target object is allowed to vary, and show that in this case the luminance can be estimated directly from the photoreceptor response. We then progressively increase the complexity of the spectral conditions and explain the correspondingly complex computations that are required to estimate the target LRV. Next, we show the optimal AMA receptive fields for each one of the spectral cases discussed in this work. We show that the optimal receptive fields for target luminance estimation have a center surround structure and they give more weights to the response of L and M cones than S cones. We finally show that, as expected, the receptive fields for the complex spectral cases generalize to the simple cases. Hence, one requires only one set of optimal receptive fields for estimating the target luminance, not separate receptive fields for each spectral case.

\subsection{Luminance estimates for each spectral case}
\subsubsection{Case 1: Target object relative surface reflectance spectrum variable, light source and background object spectra fixed}
We start with the results of the simplest type of spectral variation where only the target object relative surface reflectance spectrum is allowed to change. 
We used AMA to learn a set of linear receptive fields that are optimal for estimating target LRV from the model cone photoreceptor response to the images from this case.
Fig.~\ref{fig:case1RFResponse} shows the projection of the photoreceptor response along the first two optimal AMA receptive fields. 
Clearly, there exists a subspace along which the photoreceptor response separates according to the luminance level assigned to the target object. 
Thus, the target LRV can be obtained by projecting the photoreceptor response along these AMA receptive fields. 
We used such projections and a generalized Bayes' decoder (see methods) to estimate the luminance of the target objects in the test set stimuli. 
Fig.~\ref{fig:case9Results} shows the estimated luminance v/s the actual luminance assigned to the target object for the test stimuli in case 1. 
We show the estimates obtained by AMA and two null models: a naive method that assigns the average target LRV across the training set to all images, and a linear model based on the photoreceptors corresponding to the center pixel on the target object (see Methods). 
As expected, in the absence of any external source of variation, the photoreceptor response is proportional to the luminance of the target object, and hence can be easily estimated by both AMA and linear regression.

@@

\subsubsection{Case 2: Target object relative surface reflectance spectrum variable, light source spectra fixed, background object spectra variable}
When the light source emission spectrum is fixed, the naive expectation would be that the luminance of the target object would depend only on the surface reflectance of the target object. 
Indeed, it is believed that the variation in the surface reflectance of the background has little effect on the reflected light from the target object  \cite{BrainardWandellRetinex}. 
But our analysis shows that the secondary reflections from the target object, of the light coming indirectly after being reflected from the background objects, have a significant effect on the response of the photoreceptors corresponding to the target object. 
Fig.~{\ref{fig:case2RFResponse}} shows the projection of the photoreceptor response for the stimuli in case 2 along the first two AMA receptive fields. 
These receptive fields were learnt using the stimuli from the same case 2. 
The response projected along the first two optimal receptive fields do not separate as well compared to case 1. 
But the overlap is lower if the projection is considered in subspaces that have more AMA receptive fields. 
Fig.~\ref{fig:ErrorVsNFilters} shows the relative error of luminance estimate as a function of the number of AMA receptive fields used for luminance estimation. 
As expected, the relative error decreases with the number of receptive fields and converges to the minimum value around 8 receptive fields. 
Fig.~\ref{fig:case2Estimates} shows the luminance estimates for the learning methods used in this work. 
The mean estimates fall along the diagonal x=y line. 
The variance in the estimates increases with assigned luminance. 
This is because the isomerization noise is directly proportional to the light reflected from the target object and is higher at higher values of the target LRV.

%% Figure Case 2
\begin{figure}
\centering
    \begin{subfigure}[b]{0.3	 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure10/Figure10_a.pdf}
        \caption{RF response Case 2}
        \label{fig:case2RFResponse}
    \end{subfigure}
        \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure10/Figure10_b.pdf}
        \caption{Luminance Estimates}
        \label{fig:case2Estimates}
    \end{subfigure}
    \caption{{\bf Case 2:} (a) Photoreceptor response projected along the first two AMA receptive fields. (b) Estimated v/s assigned target LRV for the image in case 2. Solid lines show the mean estimate, the filled region in light color  shows one standard deviation.}
\label{fig:case2Results}
\end{figure}

@@

\subsubsection{Case 3: Target object relative surface reflectance spectrum variable, light source spectra variable, background object spectra fixed}
For the spectral variations of case 3, the AMA receptive fields learnt on the photoreceptor response to the images does a very poor job in predicting the target LRV (Fig.~\ref{fig:isomerizationFails}). 
This is because, since the illuminant spectra varies in this case, the photoreceptor response is not directly proportional to the target object surface anymore. 
But, since the background surfaces are fixed in this case, the ratio between the response of  photoreceptors corresponding to the target object and the background remains proportional to the target LRV. 
This is because a change in illuminant spectra produces the same effect in both the target object and the background object. 
Thus, the AMA receptive fields learnt on the contrast normalized photoreceptor response (see Methods) can be used to predict the target LRV (Fig.~\ref{fig:constrastWorks}). 
Fig.~\ref{fig:case10Results} shows the estimated target LRV v/s the assigned luminance for the images in case 3 obtained using the contrast normalized representation. 
AMA receptive fields provide excellent predictions of the target LRV. 
The predictions are much better compared to that of a linear model that only considers the (contrast normalized) response of the photoreceptors corresponding to the center pixel of the target object.

%% Figure Case 3: Isomerization Fails
\begin{figure}
\centering
    \begin{subfigure}[b]{0.3	 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure12/Figure12_a.pdf}
        \caption{RFs learnt using photoreceptor response}
        \label{fig:isomerizationFails}
    \end{subfigure}
        \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure12/Figure12_b.pdf}
        \caption{RFs learnt using contrast signal}
        \label{fig:constrastWorks}
    \end{subfigure}
            \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure12/Figure12_c.pdf}
        \caption{Luminance estimates}
        \label{fig:case10Results}
    \end{subfigure}
    \caption{{\bf Case 3: Luminance can be estimated using the contrast signal:} (a) Photoreceptor response projected on AMA receptive fields calculated using photoreceptor response. (b) The contrast normalized photoreceptor response projected along the first two AMA receptive fields. The receptive fields were calculated using the contrast normalized stimuli. The response at each luminance level separates much better in this representation. (c) Estimated v/s assigned target LRV for the image in case 3. Solid lines show the mean estimate, the filled region in light color  shows one standard deviation. AMA estimates are much better compared to the null models.}
\label{fig:importanceOfConstrast}
\end{figure}

@@

\subsubsection{Case 4: Target object relative surface reflectance spectrum variable, light source spectra variable, background object spectra variable}
% Figure Case 4
\begin{figure}
\centering
    \begin{subfigure}[b]{0.3 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure13/Figure13_a.pdf}
        \caption{AMA RF Response}
        \label{fig:case12FiltersResponse}
    \end{subfigure}    
        \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure13/Figure13_b.pdf}
        \caption{Luminance Estimates}
        \label{fig:case12Results}
    \end{subfigure}
    \caption{{\bf Case 3:} (a) Contrast normalized photoreceptor response projected along the first two AMA receptive fields. (b) Target LRV estimates v/s assigned luminance for images in case 3. The AMA estimates are unbiased and have smaller relative error compared to linear model.}
    \label{fig:case12AllResults}
\end{figure}

Case 4 is the most complex type of variation that we consider in this work. Here we allow all three types of spectra to change from image to image. In this case, neither the photoreceptor response (as in Case 1 and 2), nor any simple transformation of the response (case 3) can provide the target LRV.We used AMA to learn the receptive fields using the contrast normalized cone photoreceptor response from stimuli in case 4. Fig.~\ref{fig:case12AllResults} shows the AMA receptive field response and the luminance estimates for the images in case 4. 
We have used the first 8 receptive fields to estimate the luminance in this case (see Fig.\ref{fig:ErrorVsNFilters}). 
The AMA receptive fields do an excellent job of predicting the target LRV. 
The estimates are unbiased with the mean estimates lying close the diagonal unit slope line (Fig.~\ref{fig:case12Results}). 
The relative error of estimation is $\sim15\%$, which low compared to that of the naive model ($\sim42\%$) or the linear model on center pixel ($\sim21\%$).
